---
output: 
  pdf_document:
    latex_engine: xelatex
  
---
# Sampling
This file includes workflow and important information about preparing the dataset.
## Packages
- Tidyverse
- Data.table
```{r}
required_packages <- c("tidyverse", "data.table", "here")

for (pkg in required_packages) {
	if (!requireNamespace(pkg, quietly = TRUE)) {
		suppressMessages(install.packages(pkg))
	}
}
#load all the dependencies
invisible(lapply(required_packages, function(pkg) {
	suppressPackageStartupMessages(library(pkg, character.only = TRUE))
}))
```


## Loading the raw data
For the sampling, we need:
- review.rds
- business.rds

```{r}
#loading data.tables into environment
review <- readRDS(here("data", "raw_data","review.rds"))
business <- readRDS(here("data", "raw_data","business.rds"))

#collect $business_id vector for all Restaurant businesses
restaurant_ids <- business[
  grepl("Restaurants", categories, fixed = TRUE), 
  unique(business_id)
]

#set index to speed up the data join
setindex(review, business_id)

#data join
review_restaurants <- review[.(restaurant_ids), on = "business_id", nomatch = 0L]

#check
uniqueN(review_restaurants$business_id)  # should return 52268
```
## Counting # of reviews per Restaurant
To make sure that we have enough training data for our prediction model, a restaurant needs to have at least 50 reviews in the dataset. This makes it a necessity to compute the # of reviews per restaurant in the data.
```{r}
#stores data.table that includes business_id and corresponding number of reviews
review_counts <- review_restaurants[, .(n_reviews = .N), by = business_id]
```

## Collecting open/closed variable
Since our research is also focused on threat detection, it is important to check what percentage of restaurants is actually closed in the dataset. If this percentage is very low, we need to inflate the percentage of closed restaurants in our dataset as this will improve the model's ability to identify at-risk businesses due to the increased representation of the minority class. Without rebalancing, the model could become very "good" by predicting all businesses remain open, while completely failing to capture characteristics that could predict risk of closing.
```{r}
restaurant_status <- business[.(restaurant_ids), on = "business_id", .(business_id, is_open)]

```
The open vs. closed ratio of restaurants in the Yelp dataset is 
**`r round(mean(restaurant_status$is_open) * 100, 1)`%** open and 
**`r round((1 - mean(restaurant_status$is_open)) * 100, 1)`%** closed.

```{r}
# merge counts with open/closed status
review_counts_status <- review_counts[
  business[, .(business_id, is_open)], 
  on = "business_id"
]

# how many restaurants are closed and have >= 50 reviews?
closed_50plus <- review_counts_status[n_reviews >= 50 & is_open == 0, .N]
closed_50plus   #no of restaurants that are closed but have 50+ reviews
```
## Creating business_id vectors
In this part, we will create business_id vectors with the following characteristics:

**closed_ids vector:**
- Type = Restaurant
- Restaurant is currently closed
- No. of reviews since 01 / 01 / 2018 is #100 or higher

**open_ids vector:**
- Type = Restaurant
- Restaurant is currently open
- No. of reviews since 01 / 01 / 2018 is #100 or higher

```{r}
date_cutoff <- "2018-01-01"

    #CLOSED RESTAURANTS

closed_ids <- restaurant_status[is_open == 0, unique(business_id)]  #closed restaurants id's

since2018_closed <- review_restaurants[
  business_id %chin% closed_ids & date >= as.IDate(date_cutoff),
  .N,
  by = business_id
]   #list of restaurants that are closed and their no. of reviews since 01 / 01 / 2018

since2018_closed[N >= 100, .N]  #no. of closed restaurants that have 100+ reviews since 2018 

closed_100plus_since18 <- since2018_closed[N >= 100, business_id] #business_id vector

    #OPEN RESTAURANTS

open_ids <- restaurant_status[is_open == 1, unique(business_id)]  ##open restaurants id's

since2018_open <- review_restaurants[
  business_id %chin% open_ids & date >= as.IDate(date_cutoff),
  .N,
  by = business_id
]   #list of restaurants that are open and their no. of reviews since 01 / 01 / 2018

since2018_open[N >= 100, .N]  #no. of open restaurants that have 100+ reviews since 2018 

open_100plus_since18 <- since2018_open[N >= 100, business_id] #business_id vector

```

## Randomly selecting business_id's from the business_id vectors

We just created the following two business_id vectors:

- **closed_100plus_since18**
- **open_100plus_since18**

As discussed earlier, we will inflate the no. of closed businesses to improve the models ability to identify at-risk businesses. The percentage of closed businesses in our training dataset will be 50%.

```{r}
set.seed(2310)    #setting seed for reproducibility

sampled_closed_ids <- sample(closed_100plus_since18, 50)  #sampled business id's for closed restaurants

sampled_open_ids <- sample(open_100plus_since18, 50)      #sampled business id's for open restaurants

```

## Randomly selecting reviews for the selected restaurants

For the randomly selected business id's, we need to select a certain no. of reviews as well. We opted for 50 reviews per restaurant. This means our final training dataset will contain:

( 50 closed restaurants * 50 reviews ) + ( 50 open restaurants * 50 reviews ) = #5000 reviews.

```{r}
sampled_ids <- c(sampled_closed_ids, sampled_open_ids)  #vector of all the sampled id's

review_restaurants[, date_id := as.IDate(date)] #creating date id (YYYY - MM - DD)

review_restaurants_since18 <- review_restaurants[date_id >= as.IDate(date_cutoff)] #filtering date

setindex(review_restaurants_since18, business_id) #improves speed 

set.seed(2310)  #set seed for reproducibility

reviews_sampled <- review_restaurants_since18[
  business_id %chin% sampled_ids,   #select only rows where business_id matches with samples_ids
  .SD[sample(.N, 50)],              #sample 50 out of total .N ->
  by = business_id                  #per business_id
]

```

## Storing the training data for future use
We now store the selected traning data for future use in a training_data folder
- Check if training_data folder is present
- Create if needed
- Store a .rds file of the training data in that folder.

```{r}
# Ensure training_data directory
dir.create(here("data", "training_data"), recursive = TRUE, showWarnings = FALSE)

# Store the training data as .rds file
saveRDS(reviews_sampled, here("data","training_data","reviews_sampled.rds"))
```

